Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology

### 2014-07-31 paper.tex
Started discussion of experimental work

\documentclass[10pt,final,journal]{IEEEtran}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

\title{Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology}

@@

    Streaming cells are commonly used in membrane science to determine the zeta potential of a membrane surface.
    This is an important factor in the design and operation of membrane based processes. \cite{Daiguji2004b}
    A streaming cell contains one or more narrow channels (sometimes a porous material is used as the channel) across which a pressure gradient is applied.
    The application of a pressure gradient across the cell forces fluid through the cell and causes a streaming current, which in turn, creates voltage differential across the cell.
    This voltage differential is called the streaming potential.

@@

    Experimental results show conversion efficiencies in the range of:
    \begin{itemize}
        \item 0.01\% by forcing water through porous glass with pore sizes from 10--16\thinspace$\mu m$.\cite{Yang2003}
        \item 0.8\% by forcing pure water through a ceramic rod populated with 6\thinspace$\mu m$ pores. \cite{Yang2004}
        \item 3\% by forcing sodium chloride solution through a 75\thinspace$nm$ by 50\thinspace$\mu m$ silica channel. \cite{Heyden2007}
        \item 0.77\% by forcing sodium chloride through a 200\thinspace$nm$ pore in an alumina membrane. \cite{Lu2006}
        \item 5\% by forcing sodium chloride solution through a 0.5\thinspace$nm$ cylindrical pore in polyethylene terephalate foil. \cite{Xie2008}
    \end{itemize}

@@

    The remainder of this work is presented as follows.
    In section \ref{sect:streamingCell} we build and measure a simple streaming potential cell.
    We confirm the relationship between pressure and applied pressure and investigate the effect of channel height.
    In section \ref{sect:waterConsumption} we estimate a typical domestic water usage profile.
    This consumption profile will be used as the basis for estimating the amount of energy available to the harvester.
    Section \ref{sect:powerRequirements} determines the amount of energy an electronic meter would need to be useful.
    Following on this, section \ref{sect:harvesterSize} presents the required dimensions a usable harvester would need.
    And finally, we discuss the findings in the conclusion.

    \section{Readily Attainable Streaming Cell Power}
    \label{sect:streamingCell}
    \section{Estimation of Typical Household Water Consumption}
    \label{sect:waterConsumption}
    \section{Estimation of Power Requirements}
    \label{sect:powerRequirements}
    \section{Estimation of Streaming Cell Harvester}
    \label{sect:harvesterSize}
    \section{Conclusion}
    \bibliographystyle{plain}
    \bibliography{library}

>>>>

\documentclass[10pt,final,journal]{IEEEtran}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{siunitx}
\title{Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology}

@@

    Streaming cells are commonly used in membrane science to determine the zeta potential of a membrane surface.
    This is an important factor in the design and operation of membrane based processes.\cite{Daiguji2004b}
    A streaming cell contains one or more narrow channels (sometimes a porous material is used as the channel) across which a pressure gradient is applied.
    The application of a pressure gradient across the cell forces fluid through the cell and causes a streaming current, which in turn, creates voltage differential across the cell.
    This voltage differential is called the streaming potential.

@@

    Experimental results show conversion efficiencies in the range of:
    \begin{itemize}
        \item 0.01\% by forcing water through porous glass with pore sizes from 10\thinspace--\SI{16}{\micro\metre}.\cite{Yang2003}
        \item 0.8\% by forcing pure water through a ceramic rod populated with \SI{6}{\micro\metre} pores. \cite{Yang2004}
        \item 3\% by forcing sodium chloride solution through a \SI{75}{\nano\metre} by \SI{50}{\micro\metre} silica channel. \cite{Heyden2007}
        \item 0.77\% by forcing sodium chloride through a \SI{200}{\nano\metre} pore in an alumina membrane. \cite{Lu2006}
        \item 5\% by forcing sodium chloride solution through a \SI{0.5}{\nano\metre} cylindrical pore in polyethylene terephalate foil. \cite{Xie2008}
    \end{itemize}

@@

    The remainder of this work is presented as follows.
    In section \ref{sect:streamingCell} we build and measure a simple streaming potential cell.
    We confirm the relationship between pressure and applied pressure and investigate the effect of channel height.
    In section \ref{sect:waterConsumption} we estimate a typical domestic water usage profile.
    This consumption profile will be used as the basis for estimating the amount of energy available to the harvester.
    Section \ref{sect:powerRequirements} determines the amount of energy an electronic meter would need to be useful.
    Following on this, section \ref{sect:harvesterSize} presents the required dimensions a usable harvester would need.
    Finally we discuss the feasibility of using  in the conclusion.

    \section{Readily Attainable Streaming Cell Power} \label{sect:streamingCell}
    In \cite{Gu2000}, the authors use a relatively simple parallel plate design to generate streaming potentials.
    Sandwiching plastic shims between plates of glass provided a simple way set the channel height.
    Using this technique, the authors of \cite{Gu2000} fabricated three cells with channel heights of \SI{50}{\micro\metre}, \SI{100}{\micro\metre} and \SI{150}{\micro\metre}.
    Each of the channels were \SI{3}{\centi\metre} long and had a inner width of \SI{1}{\centi\metre}.

    We replicated the channel fabrication method and approximate dimensions to produce ten streaming cell channels found in \cite{Gu2000}.
    The channel height of our fabricated cells range from \SI{25}{\micro\metre} to \SI{250}{\micro\metre}.
    The channels have a length of \SI{3}{\centi\metre} and an internal width of \SI{1}{\centi\metre}.
    \begin{figure}
        \begin{center}
        \includegraphics[width=\linewidth]{Photo_streamingPotential_Assembly_Step3.JPG}
        \end{center}
        \caption{One of ten constructed streaming cells mounted in an acrylic holder. Copper wires have been epoxied into each side to provide a means of measuring the developed potential across the cell. The shim is visible as the two orange strips at each side of the cell.}
        \label{fig:cell}
    \end{figure}
    A photo of one of the cells is shown in Fig. \ref{fig:cell}.

    \section{Estimation of Typical Household Water Consumption}
    \label{sect:waterConsumption}
    \section{Estimation of Power Requirements}
    \label{sect:powerRequirements}
    \section{Estimation of Streaming Cell Harvester}
    \label{sect:harvesterSize}
    \section{Conclusion}
    \bibliographystyle{plain}
    \bibliography{library}
